What We Eval in the Shadows

### 2021-04-12 analysis/environments.Rmd
Up-to-date environments

```{r}
envir_var <- frequent_envir %>% filter(is.na(envir), str_detect(envir, "^(?:\\.[:alpha]|[:alpha:])[[:alpha:][:digit:]_\\.]*$"))

@@

```{r}
envir_env <- E_envir %>% filter(envir_type == "environment") 
# S4 class: from pim package, uses an environment under the hood
envir_unusual <- E_envir %>% filter(envir_type %in% c("list", "integer", "double", "NULL", "pairlist", "S4"))
envir_incorrect <- E_envir %>% anti_join(envir_env) %>% anti_join(envir_unusual)

@@

env_calls <- E_envir %>% filter(envir_type %in% c("environment", "S4")) %>%
  extract(environment_class, regex = "((?:new\\+)*)caller-([[:digit:]]+)-(.*)", into = c("new_env", "hierarchy", "specific_env"), remove = FALSE, convert = TRUE) %>%
  mutate(new_env = na_if(new_env, ""), specific_env = na_if(specific_env, ""), new_env = str_count(new_env, fixed("new+"))) %>%
  mutate(specific_env = if_else(specific_env == "emptyR_EmptyEnv", "empty", specific_env)) %>% 
  mutate(hierarchy = hierarchy - 1) # Fix. Remove with the next run
```

@@

Glabal can happen by chance, because `parent.frame()`goes up too much and reaches the global environment. SO when is it explicitly passed?

@@

@@

>>>>

```{r}
envir_var <- frequent_envir %>% filter(envir != "NA", str_detect(envir, "^(?:\\.[:alpha]|[:alpha:])[[:alpha:][:digit:]_\\.]*$"))

@@

```{r}
envir_env <- E_envir %>% filter(envir_type %in% c("environment", "S4")) 
# S4 class: from pim package, uses an environment under the hood
envir_unusual <- E_envir %>% filter(envir_type %in% c("list", "integer", "double", "NULL", "pairlist", "S4"))
envir_incorrect <- E_envir %>% anti_join(envir_env) %>% anti_join(envir_unusual)

@@

env_calls <- E_envir %>% filter(envir_type %in% c("environment", "S4")) %>%
  extract(environment_class, regex = "((?:new\\+)*)(?:caller-(-?[[:digit:]]+)-)?(.*)", into = c("new_env", "hierarchy", "specific_env"), remove = FALSE, convert = TRUE) %>%
  mutate(new_env = na_if(new_env, ""), specific_env = na_if(specific_env, ""), new_env = str_count(new_env, fixed("new+"))) %>%
  mutate(specific_env = if_else(specific_env == "emptyR_EmptyEnv", "empty", specific_env)) 
```

@@

Glabal can happen by chance, because `parent.frame()`goes up too much and reaches the global environment. So when is it explicitly passed?

@@

And when is it passed through a variable?

```{r}
variable_global <- E_envir %>% filter(environment_class == "global") %>% semi_join(envir_var)

nb_variable_global_sites <- variable_global %>% pull(src_ref) %>% n_distinct()

overview_table(
  r("nb variable global sites", nb_variable_global_sites),
  r("nb variable global site percent", ratio(nb_variable_global_sites, nb_call_sites))
)
```

So we can deduce an approximate (upper-bound) of the number of sites that touch eval not by chance!

@@

## Namespaces

In R, namespaces are different from the package environment.

```{r}
namespace_envs <- E_envir %>% filter(!str_detect(environment_class, fixed("caller-")), !str_detect(environment_class, "global|empty|base|loop"), !str_detect(environment_class, fixed("package:")))

nb_namespace_env_sites <- namespace_envs %>% pull(src_ref) %>% n_distinct()

overview_table(
  r("nb namespace env sites", nb_namespace_env_sites),
  r("nb namespace env site percent", ratio(nb_namespace_env_sites, nb_call_sites))
)
```

What are the most frequent namespaces?

```{r}
namespace_envs %>% count(environment_class, wt = duplicates, sort = TRUE) %>%
  mutate(pc = ratio(n, sum(n)), cpc = cumsum(pc))

namespace_envs %>% group_by(environment_class) %>% 
  summarise(n = n_distinct(src_ref)) %>% 
  arrange(desc(n)) %>% 
   mutate(pc = ratio(n, sum(n)), cpc = cumsum(pc))
```

### 2021-04-12 paper/main.tex
Environemnt high-level classification
Up-to-date environments

>>>>

A more high-level classification: Table~\ref{tab:highlevelenvironments}

\begin{table}[h]
    \centering
\begin{tabular}{|c|c|c|}
    \hline
    Environment & \#sites & \%sites \\
    \hline
    Parent frames &  \packageNbCallerEnvSites &  \packageCallerEnvSitePercent\\
    Global &  \packageNbGlobalEnvSites & \packageNbGlobalEnvSitePercent \\
    New & \packageNbNewEnvSites & \packageNewEnvSitePercent \\
    Base &  \packageNbBaseEnvSites & \packageNbBaseEnvSitePercent \\
    Package & \packageNbPackageEnvSites & \packagePackageEnvSitePercent \\
    Namespace & \packageNbNamespaceEnvSites & \packageNbNamespaceEnvSitePercent \\
    List &  \packageNbEnvListSites & \packageNbEnvListSitePercent \\
    Integer & \packageNbIntegerEnvirSites  & \packageNbIntegerEnvirSitePercent \\
    Null &  \packageNbNullEnvirSites & \packageNbNullEnvirSitePercent \\
    Empty & \packageNbEmptyEnvSites & \packageNbEmptyEnvSitePercent \\
    \hline
\end{tabular} \label{tab:highlevelenvironments}
\caption{Classification of environments per site.}
\end{table}
